### 2018-01-25 paper.tex
First rough rough rough sketch of paper outline

%% For double-blind review submission
\documentclass[acmlarge,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission
%\documentclass[acmlarge,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission
%\documentclass[acmlarge]{acmart}\settopmatter{}

%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format should change 'acmlarge' to
%% 'sigplan,10pt'.

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

\begin{document}

%% Title information
\title[Short Title]{Meta Analysis for Optimal Run-time Checking}         %% [Short Title] is optional;
                                        %% when present, will be used in
                                        %% header instead of Full Title.
\titlenote{with title note}             %% \titlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'
\subtitle{Subtitle}                     %% \subtitle is optional
\subtitlenote{with subtitle note}       %% \subtitlenote is optional;
                                        %% can be repeated if necessary;
                                        %% contents suppressed with 'anonymous'

%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Paper note
%% The \thanks command may be used to create a "paper note" ---
%% similar to a title note or an author note, but not explicitly
%% associated with a particular element.  It will appear immediately
%% above the permission/copyright statement.
\thanks{with paper note}                %% \thanks is optional
                                        %% can be repeated if necesary
                                        %% contents suppressed with 'anonymous'

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
Text of abstract \ldots.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{keyword1, keyword2, keyword3}  %% \keywords is optional

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}

Static analysis has several common failure modes where it is unable to discharge important safety and correctness assertions about programs. Common behaviors such as complicated loops, randomness/(stochastic behavior), and dynamic data loads (external interaction) frequently stymie static analyses. On the other hand, verifying these assertions via runtime checks requires a programmer to correctly reason about how and when to perform these checks, which can be difficult. We propose a middle-ground approach which performs \textit{online} static analyses to verify programs, and propose a tool which determines the optimal location to discharge assertions as early as possible.

\section{Motivating Examples}

Interval analysis (w/common static failure modes)
Anderson's analysis ("")
Typechecking ("", ref. GPCE)

\section{Approach: meta-analysis}

Req: program, programmer-provided annotation (to be discharged), analysis to be used to discharge the annotation, semantics for stepping program
Tool:
    Run analysis at each step, identify stoplight for annotation (red/yellow/green)
        Modify analysis to handle running in-program? (several options)
    Identify steps where stoplight result \textit{changes}
    Return to user: location of optimal check, or perhaps "can't verify"

\section{Worked examples}

go back through motivating examples

\section{Future work/Extensions}

\textit{Incremental computation} fixes the efficiency problem, move the model from pre-production checking/program modification to in-production online verification

\section{Conclusion}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
                                        %% contents suppressed with 'anonymous'
  %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
  %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
  %% acknowledge financial support and will be used by metadata
  %% extraction tools.
  This material is based upon work supported by the
  \grantsponsor{GS100000001}{National Science
    Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
  No.~\grantnum{GS100000001}{nnnnnnn} and Grant
  No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
  conclusions or recommendations expressed in this material are those
  of the author and do not necessarily reflect the views of the
  National Science Foundation.
\end{acks}

%% Bibliography
%\bibliography{bibfile}

%% Appendix
\appendix
\section{Appendix}

Text of appendix \ldots

\end{document}

### 2018-01-26 paper.tex
brainstorming in 1-1 with Jared

@@
 %% For double-blind review submission
-\documentclass[acmlarge,review,anonymous]{acmart}\settopmatter{printfolios=true}
+\documentclass[acmlarge,anonymous]{acmart}\settopmatter{printfolios=true}
 %% For single-blind review submission
@@
                         %% http://ctan.org/pkg/subcaption
+
+\usepackage{listings}
 
@@
 Req: program, programmer-provided annotation (to be discharged), analysis to be used to discharge the annotation, semantics for stepping program
+
 Tool:
+
 Run analysis at each step, identify stoplight for annotation (red/yellow/green)
+
 Modify analysis to handle running in-program? (several options)
+
 Identify steps where stoplight result \textit{changes}
+
 Return to user: location of optimal check, or perhaps "can't verify"
@@
 go back through motivating examples
+
+\begin{lstlisting}
+1 a := new int[10];
+2 r := rand();
+3 if r < magicval then
+4   x := 5
+5 else
+6   x := 15
+7 return a[x]    
+\end{lstlisting}
+\begin{tabular}{l|l|l}
+  \multicolumn{3}{c}{\textbf{Analysis results for x at entry to line 7 }}
+  \\
+  Concrete execution & (Offline) Meta Analysis & Online Analysis
+  \\
+  \hline 
+  1 exit & undet          & [5, 15]
+  \\
+  2 exit & det / redgreen & [5,5] \emph{WORLD-OR} [15,15]
+  \\
+  3 exit & det            &
+  \\
+  4 exit & $\vdots$       
+  \\
+  5 exit
+  \\
+  6 exit
+  \\
+  7 exit
+\end{tabular}
+
+Standard interval analysis result for x at entry to 7 is [5, 15].
 

### 2018-09-21 examples/binary_trees/generator.py
Add flix examples to paper repo

@@ -1,0 +1,11 @@
+def make_tree(depth,prefix,output):
+    if depth == 0:
+        return
+    for choice in ["0","1"]:
+        output.write("Edge(\"" + prefix + "\", \"" + prefix + choice + "\").")
+        make_tree(depth-1, prefix+choice, output)
+               
+for i in [2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12]:
+    print i 
+    with open("tree" + str(i) + ".flix", "w+") as f:
+        make_tree(i,"node",f)

### 2018-09-21 examples/cycles/generator.py

@@ -1,0 +1,8 @@
+def make_cycle(n,output):
+    for i in range(n):
+        i_prime = (i+1) % n
+        output.write("Edge(\"node" + str(i) + "\", \"node" + str(i_prime) + "\").")
+               
+for i in [10,25,50,75,100,125,150,175,200,225,250]:
+    with open("cycle" + str(i) + ".flix", "w+") as f:
+        make_cycle(i,f)

### 2018-09-24 paper.tex
new paper.tex doc

--- 2018-01-25 paper.tex
@@
 %% Title information
-\title[Short Title]{Meta Analysis for Optimal Run-time Checking}         %% [Short Title] is optional;
+\title[Short Title]{Patchwork: Incremental Computation for Static Analysis}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 
-Static analysis has several common failure modes where it is unable to discharge important safety and correctness assertions about programs.
-Common behaviors such as complicated loops, randomness/(stochastic behavior), and dynamic data loads (external interaction) frequently stymie static analyses.
-On the other hand, verifying these assertions via runtime checks requires a programmer to correctly reason about how and when to perform these checks, which can be difficult.
-We propose a middle-ground approach which performs \textit{online} static analyses to verify programs, and propose a tool which determines the optimal location to discharge assertions as early as possible.
+Programmers who use static analyses in their normal workflow will frequently demand the results of these analyses multiple times.
+For example, when working on code in an IDE, a programmer may make multiple small edits and request the results of an anlysis after each one.
+Since these edits are likely to be minor, performing the entire analysis from scratch each time is inefficient.
+This is a prime application for \textit{incremental computation} (insert typical IC story).
+
+Incremental computation is reasonably well understood for standard Datalog, which can express a small class of analyses.
+However, more expressiveness is needed to implement complex lattice-based analyses, particularly on infinite-height lattices, such as an interval analysis.
+We present the Patchwork framework, which extends the Datalog-based approach to allow for incremental implementations of general static analyses.
 
@@
 
-Interval analysis (w/common static failure modes)
-Anderson's analysis ("")
-Typechecking ("", ref. GPCE)
+Parity analysis (requires lattice)\\
+Interval analysis (requires widening)
 
-\section{Approach: meta-analysis}
+\section{Approach: fungilang + lattice?}
 
-Req: program, programmer-provided annotation (to be discharged), analysis to be used to discharge the annotation, semantics for stepping program
-Tool:
-    Run analysis at each step, identify stoplight for annotation (red/yellow/green)
-        Modify analysis to handle running in-program? (several options)
-    Identify steps where stoplight result \textit{changes}
-    Return to user: location of optimal check, or perhaps "can't verify"
+Present fungilang semantics
 
@@
 
-go back through motivating examples
+go back through motivating examples: parity analysis
+
+\begin{lstlisting}
+1 x = 1;
+2 while x < rand()
+3   x++;
+\end{lstlisting}
+
+\section{Evaluation}
+
+We evaluated our approach on a spread of analyses and target programs.
+
+Analyses:\\
+Simple interval? Points-to? Interprocedural?
+
+Programs:\\
+IncA benchmarks? Flix benchmarks? Both?
+
+Edit model:\\
+Adopt IncA approach? Commit history? Other?
+
+We compare to {Flix/IncA}?
 
@@
 
-\textit{Incremental computation} fixes the efficiency problem, move the model from pre-production checking/program modification to in-production online verification
+Interprocedural? Or doing that now?
+
+\section{Related Work}
+
+Incremental Datalog framework (which?):
+
+Flix: the Patchwork framework extends the work of the Flix authors.
+Specifically, Flix provides a Datalog extension with the ability to reason about lattices.
+Patchwork extends work done on incremental Datalog evaluation, to, effectively, incremental Flix evaluation.
+Notably, Flix evaluates (large numbers) of rules in the course of (example).
+This includes significant redundancy, due to (facts about their dependency analysis).
+Our approach automatically tracks these dependencies and improves performance by (numbers).
+
+IncA: the IncA papers gives an algorithm for performing incremental evaluation of lattice-aware Datalog-style programs.
+The approach takes advantage of properties of monotonic operators on lattices to quickly update aggregations and Datalog relations.
+However, the monotonicity restriction rules out important program analyses, such as (those that depend on widening).
+The IncA framework cannot express our second worked example.
+Additionally, (performance).
 
 \section{Conclusion}
+
+Existing state-of-the-art approaches to incremental program analysis are not fully general.
+Incremental Datalog engines are restricted to the simple declarative model, and even more complex approaches such as IncA restrict the types of operators an analysis user can work with.
+The Patchwork framework provides a fully general (is this true?) approach to incremental static analysis.
+Our use of Adapton also improves efficiency by (numbers).
+This represents a significant advance in the state of incremental static analysis.
 

### 2018-09-27 paper.tex
related work comparison table
Add proposed overview example for further discussion

@@
 \usepackage{listings}
+
+\usepackage{hyperref}
+\hypersetup{colorlinks = true,linkcolor = blue, urlcolor=blue}
 
@@
 
+%\newcommand{\TabYes}{\ensuremath{\ding{52}}}
+%\newcommand{\TabNo}{\ensuremath{\ding{53}}}
+\newcommand{\TabYes}{\ensuremath{\blacksquare}}
+\newcommand{\TabNo}{\ensuremath{\square}}
+
+\begin{table}
+\begin{tabular}{|p{1.3in}p{1.5in}||ccc|l|}
+  \hline
+  \textbf{program domain} &
+  \textbf{system feature} & Flix & IncA & Datafun & \emph{this paper}
+  \\
+  \hline
+  general &
+  first-class  
+  fixed points & \TabNo & \TabNo & \TabYes & \TabYes
+  \\
+  general &
+  user-defined
+  datatypes & \TabYes & ? & \TabYes & \TabYes
+  \\
+  \hline
+  datalog &
+  datalog engine & \TabYes & \TabYes & \TabYes & \TabYes~(definable)
+  \\
+  \hline
+  program analysis &
+  user-defined lattices & \TabYes & \TabYes & \TabNo & \TabYes
+  \\
+  program analysis &
+  non-monotone 
+  widening
+  & \TabYes & \TabNo & n/a & \TabYes
+  \\
+  program analysis &
+  relational 
+  analysis
+  & ? & \TabNo? & n/a & \TabYes
+  \\
+  \hline
+\end{tabular}
+\caption{By-feature comparison with related work on datalog semantics, and (incremental) datalog for program analysis.}
+\label{tab:relatedwork}
+\end{table}
+
 \section{Motivating Examples}
@@
 \end{lstlisting}
+
+\subsection{Proposed Example with Relational Domain}
+
+Consider the following simple program, which takes 1000 random walks of length 3 in $\mathbb{Z}$ and tracks how many walks end at each possible integer.
+(example taken from Min\'e \href{https://pdfs.semanticscholar.org/ccb9/2bfe24199455d7c4a430f756c915cd4e5ae8.pdf}{slides})
+\begin{lstlisting}
+1 hit = {-3:0, -1:0, 1:0, 3:0}
+2 for k = 1 to 1000:
+3   x = 0
+4   for i = 1 to 3:
+5     if random() then x++ else x--
+6   hit[x] = hit[x] + 1
+\end{lstlisting}
+
+A non-relational analysis would be unable to show that \texttt{x} is in the domain of \texttt{hit} on line 6, since such a proof depends on the relationship between $i$ and $x$ in the inner for loop.
+However,  using a relational abstract domain such as octagon, we can express the constraint that $|x|\leq i$ in the inner for loop.
+Note that this is still insufficient to show that the result will always be odd; we could, however, combine with a parity domain to show that, not only are the results in $[-3,3]$ (from octagon), they're also in $2\mathbb{Z}+1$ (from parity).
+
+We can easily vary the number of random walks taken or the length of each walk to ``grow'' the program and show asymptotic scaling.
+However, if I understood Matt's messages correctly, that doesn't address the question of an ``edit model''.
+I'll be talking to Jared about what such an edit model might look like here.
+
+\iffalse NOTE(benno): also considered a couple other examples, noting here as alternatives that could be further pursued if neeeded:
+
+This one needs relational analysis to statically discharge the assertion, but there's no good way I see initially to vary the size of the program.
+  x = 0;
+  y = 0;
+  if rand():
+    x = 1
+  else:
+    y = 1
+    assert !(x && y)
+
+This one comes directly from the original octagon analysis paper.
+It also has a nice relational story, but is similarly hard to vary the size of the program, though it is straightforward to change some initial parameters for incremental edits.
+    x = 10
+    y = 0
+    while(x > 0):
+      x--;
+      if(*):
+        y++
+    assert(x+y <= 10)  
+A non-relational analysis would not be able to statically discharge the assertion, concluding only that $x\in (-\infty,10]$ and $y\in [0,\infty)$.
+However, a relational domain like octagon is able to statically discharge the assertion using a loop invariant of $x+y\leq10\wedge-x-y\leq10$
+
+I also played around for a while with string substring operations (where we'd have to relate the length of the substring to the total length of the list and the beginning index to prove accesses safe) but couldn't come up with a clean simple example to work with.
+\end{lstlisting}
+\fi
 

### 2018-10-02 paper.tex
update to include octagon analysis example table
WIP on table showing analysis progression

@@
 \hypersetup{colorlinks = true,linkcolor = blue, urlcolor=blue}
+
+\usepackage{tikz}
+\usetikzlibrary{tikzmark}
+
+\usepackage{xcolor}
 
@@
 
+\begin{table}[]
+    \centering
+    \begin{tabular}{|c|l|c|c|c|c|}
+        \hline
+        \# & stmt & result & 1: x = 9 & 6: y += 2 & 4: x -= 2 \\
+        \hline
+        1 & x = 10; & & & &\\
+        \hline
+        2 & y = 0; & x + y $\le$ 10 & \textcolor{red}{x + y $\le$ 9} & x + y $\le$ 9 & x + y $\le$ 9\\
+        \hline
+        3 & while x > 0: & x + y $\le$ 10 & \textcolor{red}{x + y $\le$ 9} & \textcolor{red}{x + y $\le \infty$} & \textcolor{red}{x + y $\le$ 9}\\
+        \hline
+        4 & \hspace*{0.3cm} x- - & x + y $\le$ 9 & \textcolor{red}{x + y $\le$ 8} & \textcolor{red}{x + y $\le \infty$} & \textcolor{red}{x + y $\le$ 7}\\
+        \hline
+        5 & \hspace*{0.3cm} if* & & & &\\
+        \hline
+        6 & \hspace*{0.6cm} y++ & x + y $\le$ 10 & \textcolor{red}{x + y $\le$ 9} & \textcolor{red}{x + y $\le \infty$} & \textcolor{red}{x + y $\le$ 9}\\
+        \hline
+        7 & assert(x + y $\le$ 10) & true & true & \textcolor{red}{false} & \textcolor{red}{true}\\
+        \hline
+    \end{tabular}
+    \caption{Caption}
+    \label{tab:my_label}
+\end{table}
+
 \section{Motivating Examples}
@@
 I also played around for a while with string substring operations (where we'd have to relate the length of the substring to the total length of the list and the beginning index to prove accesses safe) but couldn't come up with a clean simple example to work with.
-\end{lstlisting}
+
 \fi
+
+\begin{table}[]
+  \centering
+  \def\arraystretch{2}
+    \begin{tabular}{r|l|ll}
+        \# & Statement & \multicolumn{2}{l}{Analysis State Progression} \\
+        \midrule
+        1 & \texttt{x = 10;} &$\top$\tikzmark{a} & \\
+        \hline
+        2 & \texttt{y = 0;} & $x+y\leq10$ & \\
+        \hline
+        3 & \texttt{while(x > 0)} & $x+y\leq10$ & \\
+        \hline
+        4 & \quad\texttt{x--;} &  $x+y\leq9$& \\
+        \hline
+        5 & \quad\texttt{if(*)\{} & & \tikzmark{b}Arrow testing 1 2 3\\
+        \hline
+        6 & \qquad\texttt{y++} & $x+y\leq10$& \\
+        \hline
+        7 & \quad\texttt{\}} & $x+y\leq10$ & \\
+        \hline
+%        8 & \qquad$\epsilon$ & $x+y\leq9$& \\
+%        \hline
+        8 & \texttt{assert(x + y $\le$ 10)} & $x+y\leq10$ & \\
+    \end{tabular}
+
+    \begin{tikzpicture}[overlay,remember picture, shorten >=-3pt]
+      \draw[->, red, rounded corners=8pt](pic cs:a) |- ([xshift = -5pt,yshift = 3pt]pic cs:b);
+    \end{tikzpicture}
+
+    \caption{WIP Attempt at an ``ATT'': Analysis Trace Table (tentative name)}
+    \label{tab:ATT}
+% - Table showing fixpoint computation, with a row per program location and a column per ``iterate'' (i.e.
+abstract state value at that location).
+So height of table is ~~ `n` and width is `m` in O(n*m) complexity of Tarski fixpoint: n is number of locations, m lattice height.
+Thread arrows throughout table to represent worklist causation relation - so that the order of the fixpoint computation is shown, even though the table is 2d with a ragged right edge
+\end{table}
 

### 2018-10-03 paper.tex
Finish first attempt at an analysis trace visualization

--- 2018-09-27 paper.tex
@@
 I also played around for a while with string substring operations (where we'd have to relate the length of the substring to the total length of the list and the beginning index to prove accesses safe) but couldn't come up with a clean simple example to work with.
-\end{lstlisting}
+
 \fi
+
+\begin{table}[]
+  \centering
+  \newcommand{\location}{\tikz\draw[green!65!black,fill=green!65!black] (0,0) circle (.4ex);}
+  \def\arraystretch{2}
+    \begin{tabular}{r|l|ccl}
+        \# & Program Location & \multicolumn{3}{l}{Analysis State Progression} \\
+        \midrule
+        1 & \texttt{x = 10;\location} &\quad&$\{\}$\tikzmark{1a} & \\
+        \hline
+        2 & \texttt{y = 0;\location} &$\bot$\quad& $\{x+y\leq10\}$\tikzmark{2a} & \hspace{1em}$\{x+y\leq10\}$\tikzmark{2b}\\
+        \hline
+        3 & \texttt{while(x > 0)\{\location} &$\bot$\quad& $\{x+y\leq10\}$\tikzmark{3a} & \tikzmark{3b}\\
+        \hline
+        4 & \quad\texttt{x--;\location} &$\bot$\quad&  $\{x+y\leq9\}$\tikzmark{4a}& \\
+        \hline
+        5 & \quad\texttt{if(*) \{y++;\location\}} &$\bot$\quad&$\{x+y\leq10\}$\tikzmark{5a}& \\
+        \hline
+        6 & \quad\texttt{else \{$\epsilon$;\location\}} &$\bot$\quad& $\{x+y\leq9\}$\tikzmark{6a} & \\
+        \hline
+        7 & \location\texttt{\}} &$\bot$\quad& $\{x+y\leq10\}$\tikzmark{7a} & \tikzmark{7b}\\
+        \hline
+%        8 & \qquad$\epsilon$ & $x+y\leq9$& \\
+%        \hline
+        8 & \location~\texttt{assert(x + y $\le$ 10)} &$\bot$\quad& $\{x+y\leq10\}$\tikzmark{8a} & \hspace{1em}$\{x+y\leq10\}$\tikzmark{8b}\\
+    \end{tabular}
+
+    \newcommand{\abovetext}{10pt}
+    \newcommand{\belowtext}{-4pt}
+
+    \begin{tikzpicture}[overlay,remember picture, shorten >=-3pt]
+      \coordinate (1a) at ([xshift=-0.46em]pic cs:1a) ;
+      \coordinate (2a) at (pic cs:2a);
+      \coordinate (3a) at (pic cs:3a);
+      \coordinate (4a) at (pic cs:4a);
+      \coordinate (5a) at (pic cs:5a);
+      \coordinate (6a) at (pic cs:6a);
+      \coordinate (7a) at (pic cs:7a);
+      \coordinate (8a) at (pic cs:8a);
+      \draw[->, blue!75!black] ([yshift=\belowtext]1a) -- ([yshift = \abovetext]1a|-2a);
+      \draw[->, blue!75!black] ([yshift=\belowtext]1a|-2a) -- ([yshift = \abovetext]1a|-3a);
+      \draw[->, blue!75!black] ([yshift=\belowtext]1a|-3a) -- ([yshift = \abovetext]1a|-4a);
+      \draw[->, blue!75!black] ([yshift=\belowtext]1a|-4a) -- ([yshift = \abovetext]1a|-5a);
+      \draw[->, blue!75!black] ([yshift=\belowtext]1a|-4a)
+      to [out=270,in=90] ([xshift=0.6em,yshift=0.2em] pic cs:5a)
+      to [out=270,in=90] ([yshift = \abovetext]1a|-6a);
+      \draw[->, blue!75!black] ([yshift=\belowtext]1a|-5a)
+      to [out=270,in=90] ([xshift=-5.25em,yshift=0.25em] pic cs:6a)
+      to [out=270,in=90] ([yshift = \abovetext]1a|-7a);
+      \draw[->, blue!75!black] ([yshift=\belowtext]1a|-6a) -- ([yshift = \abovetext]1a|-7a);
+      \draw[->, blue!75!black,rounded corners=4pt] ([yshift=4pt,xshift=1pt] pic cs:7a) -| ([xshift=-1pt] pic cs:3b) |- ([xshift=-5.5em, yshift=3pt]pic cs:2b);
+      \draw[->, blue!75!black,rounded corners=4pt] ([yshift=4pt,xshift=1pt] pic cs:7a) -| ([yshift=-5pt, xshift=-1pt] pic cs:7b) |- ([xshift=-5.5em, yshift=3pt]pic cs:8b);
+      \draw[->, blue!75!black,rounded corners=3pt] ([yshift=3pt,xshift=-5.25em] pic cs:2a) -| ([xshift=-5.65em,yshift=0.35em] pic cs:6a) |- ([yshift=3pt,xshift=-5.5em] pic cs:8a);
+    \end{tikzpicture}
+
+    \caption{ Visualization of fix-point computation for a program analyzer performing octagon analysis on the example program from Fig.~\ref{???}.
+      Program locations in the abstract syntax are denoted by green circles in the concrete syntax.
+The blue arrows illustrate a ``trace'' of the worklist/fix-point computation, showing the propagation of dataflow analysis results through the program.
+%      Thus, the full history of partially-computed abstract states at a particular location is recorded from left to right in the ``Analysis State Progression'' column.
+    }
+    \label{tab:ATT}
+\end{table}
 

### 2018-10-13 paper.tex
Fix analysis trace diagram by removing spurious edge
first draft of analysis loop: show calls to 'update', to 'push', and to 'pop' in the loop

@@
 
+\begin{figure}
+\begin{lstlisting}
+type Q
+type Inv
+type VisitRes = NoChange | Change Inv
+
+let visitloc : Inv -> Loc -> VisitRes =
+  lam inv loc. 
+    ...(elided program analysis logic)...
+    ...
+    ...   let inv' = update inv name loc absstate ;
+    ...   Change inv'
+    ...
+    ...(elided program analysis logic)...
+
+letrec loop : Inv -> Q -> Inv =
+  lam inv q.
+    match pop q with
+      None -> inv
+      Some q loc ->
+         match visitloc loc inv with
+           NoChange -> loop inv q
+           Change inv -> loop inv (push q (succof loc))    
+\end{lstlisting}
+\end{figure}
+
 %\newcommand{\TabYes}{\ensuremath{\ding{52}}}
@@
         \hline
-%        8 & \qquad$\epsilon$ & $x+y\leq9$& \\
-%        \hline
-        8 & \location~\texttt{assert(x + y $\le$ 10)} &$\bot$\quad& $\{x+y\leq10\}$\tikzmark{8a} & \hspace{1em}$\{x+y\leq10\}$\tikzmark{8b}\\
+        8 & \location~\texttt{assert(x + y $\le$ 10)} &$\bot$\quad& $\{x+y\leq10\}$\tikzmark{8a} & \\
     \end{tabular}
@@
       \draw[->, blue!75!black,rounded corners=4pt] ([yshift=4pt,xshift=1pt] pic cs:7a) -| ([xshift=-1pt] pic cs:3b) |- ([xshift=-5.5em, yshift=3pt]pic cs:2b);
-      \draw[->, blue!75!black,rounded corners=4pt] ([yshift=4pt,xshift=1pt] pic cs:7a) -| ([yshift=-5pt, xshift=-1pt] pic cs:7b) |- ([xshift=-5.5em, yshift=3pt]pic cs:8b);
+%      \draw[->, blue!75!black,rounded corners=4pt] ([yshift=4pt,xshift=1pt] pic cs:7a) -| ([yshift=-5pt, xshift=-1pt] pic cs:7b) |- ([xshift=-5.5em, yshift=3pt]pic cs:8b);
       \draw[->, blue!75!black,rounded corners=3pt] ([yshift=3pt,xshift=-5.25em] pic cs:2a) -| ([xshift=-5.65em,yshift=0.35em] pic cs:6a) |- ([yshift=3pt,xshift=-5.5em] pic cs:8a);

### 2018-10-14 paper.tex
add a version of the analysis loop with explicit names

@@
 \usepackage{listings}
+\lstset{}
 
@@
 \citestyle{acmauthoryear}   %% For author/year citations
+
+% Darker versions (dRed, dBlue, etc.) of common colours;
+% pure red/blue/green are violently garish, and too light in monochrome.
+%
+% dDkRed, etc. are darker still.
+% dLight[er]{Blue|Purple} are lighter colours suitable for backgrounds.
+%
+% dFaint, dGray, dDark, dAlmostBlack are shades of gray.
+%
+\definecolor{dRed}{rgb}{0.65, 0.0, 0.0}
+\definecolor{dDkRed}{rgb}{0.35, 0.0, 0.0}
+\definecolor{dGreen}{rgb}{0.0, 0.65, 0.0}
+\definecolor{dDkGreen}{rgb}{0.0, 0.35, 0.0}
+\definecolor{dBlue}{rgb}{0.0, 0.0, 0.65}
+\definecolor{dLightBlue}{rgb}{0.4, 0.4, 0.9}
+\definecolor{dLighterBlue}{rgb}{0.8, 0.8, 1.0}
+\definecolor{dDkBlue}{rgb}{0.0, 0.0, 0.45}
+\definecolor{dLightPurple}{rgb}{0.9, 0.5, 0.9}
+\definecolor{dLighterPurple}{rgb}{1.0, 0.7, 1.0}
+\definecolor{dPurple}{rgb}{0.65, 0.0, 0.65}
+\definecolor{dDigPurple}{rgb}{0.5, 0.0, 0.5}
+% \definecolor{DDIGPURPLE}{rgb}{0.5, 0.0, 0.5}  % laughable
+\definecolor{dFaint}{rgb}{0.7, 0.7, 0.7}
+\definecolor{dGray}{rgb}{0.5, 0.5, 0.5}
+\definecolor{dDark}{rgb}{0.2, 0.2, 0.2}
+\definecolor{dAlmostBlack}{rgb}{0.1, 0.1, 0.1}
+
+%%% ------- Fungi code in latex listings:
+\definecolor{darkgreen}{rgb}{0,0.5,0}
+\definecolor{darkpurple}{rgb}{0.5,0,0.5}
+\lstset{emph={Nm, Nat, Ref, Type, type, match, with, if, then, else, not, in, let, letrec, ref, get, thunk, force, scope, leaf, NmSet, Nat, Nm, idxtm, nmtm},
+  %emphstyle={\color{blue}\bfseries},
+  emphstyle={\bfseries},
+  keywordstyle={\bfseries},
+  commentstyle={\color{darkpurple}\ttfamily},
+  language=C,
+  showstringspaces=false,
+}%
+\lstset{emph={[2]%  
+    true, false,
+    },emphstyle={[2]\color{darkgreen}\bfseries}%,
+}%
+\newcommand{\code}[1]{\lstinline[basicstyle=\ttfamily]|#1|}
+\newcommand{\hlcode}[1]{{\hl{\texttt{#1}}}}
+\newcommand{\vttfamily}{\fontfamily{cmvtt}\selectfont}
+\lstset{
+%basicstyle=\fontsize{7}{8}\ttfamily,
+basicstyle=\fontsize{8}{9}\ttfamily,
+mathescape=true,
+literate={lam~}{$\lambda$}1
+         %%%%%%%%%%% Arrows
+         %%%%%%%%%%% ---------------------
+         {...}{${\color{blue}{\cdots}}$}3
+         {=>}{$\Rightarrow$}2
+         {<=}{$\leq$}2
+         {->}{$\rightarrow$}2
+         {nm->}{$\rightarrow$}2 %-- Name term function sort
+         {idx->}{$\rightarrow$}2 %-- Index term function sort
+         {idx=>}{$\Rightarrow$}2 %-- Kind arrow: sort to kind
+         %% {->}{${\color{dBlue}{\rightarrow}}$}2
+         %% {nm->}{${\color{dPurple}{\rightarrow}}$}2 %-- Name term function sort
+         %% {idx->}{${\color{dRed}{\rightarrow}}$}2 %-- Index term function sort
+         %% {idx=>}{${\color{dRed}{\Rightarrow}}$}2 %-- Kind arrow: sort to kind
+         %%%%%%%%%% Balanced delims:
+         %%%%%%%%%% --------------------------------
+         % {{ ... }} = Name set literals (index terms)
+         {<~}{${\color{dRed}{\left\{\right.}}$}1
+         {~>}{${\color{dRed}{\left.\right\}}}$}1
+         % [[ ... ]] = Annotations of types with index terms (e.g., name sets)
+         {[[}{${\color{dRed}{\left[\right.}}$}1
+         {]]}{${\color{dRed}{\left.\right]}}$}1
+         % <[ ... ]> = Annotations of names (e.g., Write scope)
+         {<[}{${\color{dRed}{\left<\right.}}$}1
+         {]>}{${\color{dRed}{\left.\right>}}$}1
+         % [< ... >] = Thunk boundaries in concrete expression syntax
+         {[<}{${\color{blue}{\left\{\right.}}$}1
+         {>]}{${\color{blue}{\left.\right\}}}$}1
+         % << ... ,, ... >> = Binary name composition; primitive form
+         {<<}{${\color{dPurple}{\left<\right.\hspace{-4pt}\left<\right.}}$}1
+         {>>}{${\color{dPurple}{\left.\right>\hspace{-4pt}\left.\right>}}$}1
+         {,,}{${\color{dPurple}{,}}$}1
+         %%%%%%%%%%
+         %%%%%%%%%% Operators, and index term syntax:
+         %%%%%%%%%% ---------------------------------
+         {~@~}{${\color{dPurple}{\bullet}}$}1 %-- Binary name composition (for Nm sort)
+         {~@@~}{${\color{dRed}{\bullet}}$}1 %-- Binary name composition (for NmSet sort)
+         {~*}{${}^{{\color{dRed}{\ast}}}$}1 %-- Kleene star; for an index term function over an initial set
+         {@}{{\textcolor{violet}{@}}}1 %-- Name literal/constant
+         {^}{${\color{dRed}{\bot}}$}1 %--- Apart symbol
+         {forall~}{${\color{dRed}{\forall}}$}1 %--- Forall index terms
+         {exists~}{${\color{dRed}{\exists}}$}1 %--- Exists index term
+         {writeset~}{${\color{dRed}{\rhd}}$}1  %--- Write set (part of a computation type)
+         %%%%%%%%% Sorts:
+         %{NmSet}{{\color{dRed}{NmSet}\color{black}}}4
+         %% {leaf}{{\color{dPurple}{leaf}\color{black}}}3
+         %% {Nm~}{{\color{dPurple}{Nm}\color{black}}}2
+         %% %{Nat~}{{\color{dRed}{Nat}\color{black}}}4
+         {Nat~}{Nat}4
+         %% %%%%%%%%%% Declarations
+         %% %%%%%%%%%% ------------------------------------
+         %% {idxtm}{{\color{dRed}{idxtm}\color{black}}}4
+         %% {nmtm}{{\color{dPurple}{nmtm}\color{black}}}4
+}
 
@@
 let visitloc : Inv -> Loc -> VisitRes =
-  lam inv loc. 
-    ...(elided program analysis logic)...
+  lam~inv.lam~loc.
     ...
+    ...   // subcase: Update the invariant mapping:
+    ...   let inv' = update inv loc absstate ;
+    ...   Change inv'
+    ...
+
+letrec loop : Inv -> Q -> Inv =
+  lam~inv.lam~q.
+    match pop q with
+      None => inv
+      Some q loc =>
+         match visitloc loc inv with
+           NoChange   => loop inv q
+           Change inv => loop inv (push q (succof loc))    
+\end{lstlisting}
+\caption{Analysis loop: Each ``visit'' to a program location has the
+  potential to \code{update} the invariant mapping \code{inv}; the
+  queue~{q} controls the schedule of these visits. }
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}
+/// Version with explicit naming strategy
+type Q[X:NmSet]
+type Inv[X:NmSet]
+type VisitRes[X:NmSet] = NoChange | Change (Inv[X])
+
+let visitloc : Inv[X] -> Loc[Y] -> exists Z|Z^X. VisitRes[X^Z] =
+  lam~inv.lam~loc.
+    ...
+    ...   // Construct a name Z such that Z^X
+    ...   let name : Nm[Z] = ... ;
     ...   let inv' = update inv name loc absstate ;
@@
     ...
-    ...(elided program analysis logic)...
 
-letrec loop : Inv -> Q -> Inv =
-  lam inv q.
+letrec loop : Inv[X] -> Q[Y] -> exists Z|Z^X. Inv[X^Z] =
+  lam~inv.lam~q.
     match pop q with
-      None -> inv
-      Some q loc ->
+      None => inv
+      Some q loc =>
          match visitloc loc inv with
-           NoChange -> loop inv q
-           Change inv -> loop inv (push q (succof loc))    
+           NoChange   => loop inv q
+           Change inv => loop inv (push q (succof loc))
 \end{lstlisting}
+\caption{Analysis loop: Each ``visit'' to a program location has the
+  potential to \code{update} the invariant mapping \code{inv}; the
+  queue~{q} controls the schedule of these visits. }
 \end{figure}

### 2018-12-10 paper.tex
update paper draft based on 12-3 discussion

@@
 Parity analysis (requires lattice)\\
-Interval analysis (requires widening)
+Interval analysis (requires widening)\\
+Octagon analysis (relational)
 
-\section{Approach: fungilang + lattice?}
+\section{Contribution: Incremental Relation Structures}
 
-Present fungilang semantics
+We developed an incremental data structure to represent the abstract states necessary to back our static analyses. It is...
+
+\section{Contribution: Extending to Incremental Analysis Framework}
+
+We use the discussed data structures in a modular framework for implementing static analyses. Our framework is organized as follows...
 
@@
 Analyses:\\
-Simple interval? Points-to? Interprocedural?
+Interval, octagon.
 
 Programs:\\
-IncA benchmarks? Flix benchmarks? Both?
+IncA benchmarks?
 
 Edit model:\\
-Adopt IncA approach? Commit history? Other?
+Adopt IncA approach?
 
-We compare to {Flix/IncA}?
+We compare to: IncA, off-the-shelf analysis?
+
+We show: same alarm rates as off-the-shelf analysis, similar incremental performance as IncA. QoIs: cumulative time, alarm rates, false alarm rates
+
+We need: off-the-shelf analysis to use, benchmarks to use, edit model to use
 

### 2019-01-18 paper.tex
Jedi-text for abstract, sentence-level intro edits, some remarks in intro

@@
 \definecolor{darkgreen}{rgb}{0,0.5,0}
+\definecolor{darkred}{rgb}{0.6,0,0}
 \definecolor{darkpurple}{rgb}{0.5,0,0.5}
@@
 \newcommand{\vttfamily}{\fontfamily{cmvtt}\selectfont}
+\newcommand{\remark}[1]{\textcolor{darkred}{$\blacktriangleright$#1$\blacktriangleleft$}}
+
 \lstset{
@@
 \begin{abstract}
-Text of abstract \ldots.
+  Proposed abstract jedi text:
+  \begin{itemize}
+  \item Program analysis, to be useful for developers, must surface results quickly during development loop.
+  \item Recomputing analysis with each edit is inefficient, performs redundant work
+  \item Existing incremental techniques express limited class of analyses
+  \item Our technique uses fungi/adapton incremental techniques to incrementally recompute analyses on-demand, (1) guaranteeing from-scratch consistency, (2) performing minimal recomputation, and (3) expressing broader class of analyses than existing techniques.
+  \item Key insight: fine grained memo-matching on analysis execution traces, adapton-style dependency graph.
+  \end{itemize}
 \end{abstract}
@@
 Programmers who use static analyses in their normal workflow will frequently demand the results of these analyses multiple times.
-For example, when working on code in an IDE, a programmer may make multiple small edits and request the results of an anlysis after each one.
-Since these edits are likely to be minor, performing the entire analysis from scratch each time is inefficient.
+For example, when working on code in an IDE, a programmer may make multiple small edits and request the results of an analysis after each one.
+However, since edits typically affect only a small region of the program, it is very inefficient to recompute a whole-program analysis with each edit.
+
 This is a prime application for \textit{incremental computation} (insert typical IC story).
 
-Incremental computation is reasonably well understood for standard Datalog, which can express a small class of analyses.
-However, more expressiveness is needed to implement complex lattice-based analyses, particularly on infinite-height lattices, such as an interval analysis.
-We present the Patchwork framework, which extends the Datalog-based approach to allow for incremental implementations of general static analyses.
+Incremental computation is reasonably well understood for standard Datalog, which can express a limited class of analyses.\cite{flix, datafun, IncA, ???} However, more expressivity is needed \remark{Why?
+we need to justify this.} to implement complex lattice-based analyses, particularly those with infinite-height abstract domain lattices, including many popular numerical (e.g.
+intervals, octagons) and logical (e.g.
+separation logic, predicate abstraction) abstract domains.
+
+We present the Patchwork framework, which extends the Datalog-based approach \remark{Are we really ``extending'' a datalog approach?
+I'd be inclined to say something like ``unlike datalog approaches, we can do ...'' or avoid mentioning datalog altogether}to allow for incremental implementations of general static analyses.
 

### 2019-01-22 paper.tex
Fill out incremental contribution sections

--- 2018-12-10 paper.tex
@@
 
-We developed an incremental data structure to represent the abstract states necessary to back our static analyses. It is...
+We developed an incremental data structure to represent the abstract states necessary to back our static analyses.
+Our representation allows for efficient incremental joins to build up the underlying relation.
+We use the following naming strategy...
+
+You are convinced (*wave hand*) that the data structures in question provides a unique performance improvement for our use case.
 
@@
 
-We use the discussed data structures in a modular framework for implementing static analyses. Our framework is organized as follows...
+We use the discussed data structures in a modular framework for implementing static analyses.
+Our framework requires the implementation of an abstract domain, including transfer function, to be given by the user.
+This approach allows us to take advantage of existing libraries used in the static analysis community.
+For this paper, we interfaced with APRON, an existing library supporting the numerical domains necessary for interval and octagon analysis.
+Once we have defined the language to operate on - in our case, Jimple - and the transfer function for our domain of choice (see examples), we can incrementally execute the analysis.
+
+See figures 1/2 for an outline of the driving functions for the incremental analysis.
+By exploiting the structure of the typical 'worklist' algorithm for static analysis, we can achieve incremental performance gains.
+
+Our naming strategy works by reference to (insert defense of naming strategy - strictly location?) (reference to analysis trace visualization - given by Benno in table 3)
 

### 2019-01-28 paper.tex
Modify abstract/intro/related work with new framing

@@
   \item Existing incremental techniques express limited class of analyses
-  \item Our technique uses fungi/adapton incremental techniques to incrementally recompute analyses on-demand, (1) guaranteeing from-scratch consistency, (2) performing minimal recomputation, and (3) expressing broader class of analyses than existing techniques.
+  \item Our technique takes an interpretative approach, avoiding the need to compile a program to a set of facts
+  \item This allows us to express a broader set of analyses over user-defined languages
   \item Key insight: fine grained memo-matching on analysis execution traces, adapton-style dependency graph.
@@
 
-Incremental computation is reasonably well understood for standard Datalog, which can express a limited class of analyses.\cite{flix, datafun, IncA, ???} However, more expressivity is needed \remark{Why?
-we need to justify this.} to implement complex lattice-based analyses, particularly those with infinite-height abstract domain lattices, including many popular numerical (e.g.
-intervals, octagons) and logical (e.g.
-separation logic, predicate abstraction) abstract domains.
+Existing approaches are in what we call 'Datalog-style', wherein a program is transformed into a set of initial facts, and the analysis consists of applying a series of rules until a fixed-point is reached.
+Examples of this approach include Flix (non-incremental) and IncA (incremental).
+This approach loses access to the underlying program as the analysis runs, which poses problems for incremental computation.
+In particular, IncA cannot express analyses requiring the use of non-monotonic widening operators, which includes most that use infinite-height lattices, such as interval or \remark{compelling examples here}.
 
-We present the Patchwork framework, which extends the Datalog-based approach \remark{Are we really ``extending'' a datalog approach?
-I'd be inclined to say something like ``unlike datalog approaches, we can do ...'' or avoid mentioning datalog altogether}to allow for incremental implementations of general static analyses.
+We present the Patchwork framework, which uses Adapton's DCG-based incremental computation.
+This approach enables us to bring incremental programming's efficiency to general static analysis over arbitrary user-defined languages.
+Once the user writes a transfer function over their desired language for their desired analysis, Patchwork executes the analysis and performs any desired re-computations quickly.
 
@@
 
-Flix: the Patchwork framework extends the work of the Flix authors.
-Specifically, Flix provides a Datalog extension with the ability to reason about lattices.
-Patchwork extends work done on incremental Datalog evaluation, to, effectively, incremental Flix evaluation.
-Notably, Flix evaluates (large numbers) of rules in the course of (example).
-This includes significant redundancy, due to (facts about their dependency analysis).
-Our approach automatically tracks these dependencies and improves performance by (numbers).
+Flix: Flix enriches the basic semantics of Datalog with lattices, making static analysis significantly easier to write.
+However, Flix is still tied to the core Datalog model - compiling a program to an initial set of facts, then applying a set of productions rules to a fixpoint - which limits the efficiency of the algorithm.
+Specifically, Flix evaluates (a lot) of rules in the course of (example), including a large number of redundant evaluations, due to (facts about their dependency analysis).
+The DCG-based approach automatically tracks dependencies in a much better (why, specifically) model, which improves efficiency.
 
-IncA: the IncA papers gives an algorithm for performing incremental evaluation of lattice-aware Datalog-style programs.
-The approach takes advantage of properties of monotonic operators on lattices to quickly update aggregations and Datalog relations.
-However, the monotonicity restriction rules out important program analyses, such as (those that depend on widening).
-The IncA framework cannot express our second worked example.
-Additionally, (performance).
+IncA: IncA effectively builds on work like that in Flix by adding explicit support for incremental computation.
+After running a static analysis, the user can modify the initial set of facts (implicitly modifying the program) and re-execute with some improvements in performance.
+IncA's approach takes advantage of properties of monotonic operators on lattices to quickly update analysis results for certain types of edits.
+However, the restrictions that all operators must be monotonic rules out common and important program analysis, such as (examples from above).
+For instance, IncA's framework could not express (our worked example) - they would instead have to use (different widening operator), which would significantly degrade the strength of the analysis (for reasons).
 

### 2019-02-05 paper.tex
Add section w/detailed explanation of Datalog strategy

@@
 %% Title information
-\title[Short Title]{Patchwork: Incremental Computation for Static Analysis}         %% [Short Title] is optional;
+\title[Short Title]{Patchwork: Incremental Computation for Abstract Interpretation}         %% [Short Title] is optional;
                                         %% when present, will be used in
@@
 
-\section{Contribution: Incremental Relation Structures}
+\section{Datalog Strategy - Limitations + Comparison}
 
-We developed an incremental data structure to represent the abstract states necessary to back our static analyses.
-Our representation allows for efficient incremental joins to build up the underlying relation.
-We use the following naming strategy...
+The Datalog-style approach to static analysis uses the following broad algorithm:
 
-You are convinced (*wave hand*) that the data structures in question provides a unique performance improvement for our use case.
+\begin{enumerate}
+    \item Define a set of relations
+    \item Define a set of rules that refer to these relations
+    \item Transform the input program into a set of 'facts', which are judgements of membership in the relations
+    \item Apply the rules to generate new facts until a fixpoint is reached
+\end{enumerate}
+
+\subsection{Example}
+
+As an example, consider the simple points-to analysis defined by the Flix authors (From Datalog to FLIX: A Declarative Language for Fixed Points on Lattices, section 2).
+The approach is applied as follows:
+
+\begin{enumerate}
+    \item Define a set of relations.\\
+    For this problem, there are 6 relations total - 4 "input relations", which are the base facts, and 2 "derived relations", which give the information we're interested in.
+They are as follows: New(v, h), Assign(v, v), Store(v, f, v), Load(v, v, f) are the inputs, and VarPointsTo(v, h) and HeapPointsTo(h, f, h) are the derived relations.
+Here the lowercase letters are types - v for variable, h for heap, f for field.
+    \newpage
+
+    \item Define a set of rules that refer to these relations.\\
+    There are four rules that define this analysis:
+    \begin{enumerate}
+        \item VarPointsTo(v1, h1) :- New(v1, h1)
+        \item VarPointsTo(v1, h2) :- Assign(v1, v2), VarPointsTo(v2, h2)
+        \item VarPointsTo(v1, h2) :- Load(v1, v2, f), VarPointsTo(v2, h1), HeapPointsTo(h1, f, h2)
+        \item HeapPointsTo(h1, f, h2) :- Store(v1, f, v2), VarPointsTo(v1, h1), VarPointsTo(v2, h2)
+    \end{enumerate}
+    For each of the rules, the fact on the left is added to the database if all of the facts on the right are in the database.\\
+
+    \item Transform the input program into a set of 'facts'.
+    \begin{lstlisting}[language=Java]
+    ClassA o1 = new ClassA(); // object A
+    ClassB o2 = new ClassB(); // object B
+    ClassB o3 = o2;
+    o2.f = o1;
+    Object r = o3.f;
+    \end{lstlisting}
+
+    This program is transformed into the following facts:\\
+    New("o1", "A")\\
+    New("o2", "B")\\
+    Assign("o3", "o2")\\
+    Store("o2", "f", "o1")\\
+    Load("r", "o3", "f")\\
+
+    The algorithm for this transformation is not given.
+For this simple example, it's fairly easy to see how each fact was derived from the program.
+Some algorithm for this transformation must be written for any analysis in this style.\\
+
+    \item Apply the rules to generate new facts until a fixpoint is reached.\\
+    This results in the following derived relations:\\
+    VarPointsTo("o1", "A") - rule 1\\
+    VarPointsTo("o2", "B") - rule 1\\
+    VarPointsTo("o3", "B") - rule 2\\
+    HeapPointsTo("B", "f", "A") - rule 4\\
+    VarPointsTo("r", "A") - rule 3\\
+
+    In the example, the final fact is what was desired.
+\end{enumerate}
+
+\subsection{Naive Incremental Computation in the Datalog World}
+
+In the Datalog context, an edit consists of the insertion or deletion of facts to the starting database.
+Wlog, we consider insertions and deletions of single facts and discuss naive incremental strategies for these edits.
+
+In the course of the Datalog engine's run, new facts are added to the database when a rule is triggered.
+It is easy to construct a simple 'provenance' of each fact in the database by tagging the fact with the facts that generated it.
+For instance, in the example from 3.1, the VarPointsTo("o3", "B") was generated via the facts Assign("o3", "o2") and VarPointsTo("o2, "B"), the latter of which was itself generated from New("o2", "B").
+Assuming that in the course of the algorithm these provenances are stored, consider now an edit which consists of the deletion of the fact New("o2", "B") from the initial database.
+Rather than re-computing from scratch, we can start from our finished database and do the following: delete the fact New("o2", "B"), delete all facts that transitively depend on New("o2", "B"), use the remaining facts as the initial database for a new run.
+Crucially, rather than re-computing all the facts from scratch, using this algorithm, any computed fact which does not depend on the deleted fact does not have to be re-computed.
+
+Now consider an edit which consists of the insertion of an arbitrary fact.
+This does not affect the facts in the database, which means we can do the following: simply use the finished database + the new fact as the initial database for a new run.
+But, what if the fact adds to a relation that is negated in one of our rules?
+Then the insertion of this fact is also a deletion of the corresponding negation, and the deletion should be processed first.
+(Datalog engines typically have restrictions on the use of negation in their rules - not discussed here in detail)
+
+Therefore, a naive incremental algorithm for an arbitrary set of deletions and insertions $\Delta\{d_k, i_k\}$ is as follows: we have a final database $D$.
+For each deletion $d_k$ , delete $d_k$ and all facts that transitively depend on it from $D$, leaving us with $D'$.
+Then add each inserted fact to get $D' + \{i_k\}$, and do a new run starting with this database.
+
+\subsection{IncA's Improved Approach}
+
+The above are the simplest two incremental strategies which many Datalog engines (even those that do not advertise themselves as incremental) implement.
+IncA, an explicitly incremental framework for Datalog (+ lattices), adds an additional optimization in the context of rules over lattices.
+Assuming all functions used are \textit{monotone}, a deletion of fact $d$ and insertion of fact $i$ where $d \sqsubseteq i$ constitutes a \textit{monotonically increasing replacement} (term introduced by IncA authors).
+If all functions used are monotone, then it is true that any fact derived from $d$ is also derivable from $i$ (proof left to IncA folks).
+Therefore, rather than deleting many facts derived from $d$ and subsequently re-deriving all of them, we don't have to perform any deletions.
+
+Therefore IncA's improved incremental algorithm is as follows.
+Given a final database $D$ and an arbitrary set of deletions and insertions $\Delta\{d_k, i_k\}$, first identify which $d_c$ have corresponding $i_c$ such that $d_c \sqsubseteq i_c$.
+Delete each of these from $\{d_k\}$ to get the strictly smaller set $\{d'_k\}$.
+Then run the above naive algorithm on $\Delta\{d'_k, i_k\}$
+
+There are some additional implementation details to IncA's algorithm that let it efficiently perform these deletions/re-derivations, but those do not change the improvement that's being made in terms of incremental computation, so we elide those details here.
+
+\subsection{Limitations}
+
+Once the Datalog strategy is chosen, there are several inherent limitations with respect to incremental computation.
+We briefly outline several of them here, and gesture toward how these are overcome using the Adapton strategy.\\
+
+Both Flix and IncA require that all functions over lattices are \textit{monotone}.
+This is a non-trivial restriction in the context of static analysis, as many standard analyses use non-monotonic widening functions to ensure termination.
+All of these are inadmissible for Flix or IncA.
+Adapton has no monotonicity restriction.\\
+
+There is an important sense in which the approach taken by IncA lacks 'history' in the sense of Adapton.
+We illustrate via an example.
+Consider an arbitrary lattice with elements $a, b, c, d$ such that the following holds: $a$ and $b$ are not ordered, $a \sqcup c = d$, $b \sqcup c = d$, $c \neq d$.
+Assuming that $a$ and $c$ are in the initial set of facts, an analysis run will add $d$ to the database.
+Now assume we make the following edit: delete $a$ and insert $b$.
+This is not a monotonically increasing replacement, so IncA will transitively delete all facts depending on $a$, including $d$ and any facts derived from $d$.
+Now after adding $b$, in the incremental run, the analysis will quickly re-discover $d$.
+However, since $d$ was already deleted from the database, all consequents of $d$ must be completely re-computed.
+Because IncA cannot 'look ahead' to see that $d$ will be re-discovered from an inserted fact, work that could have been saved is lost.
+This is a specific consequence of the general problem that deleting all facts deriving from a deleted fact is an over-approximation that is necessary for the Datalog approach to be sound.
+In contrast, Adapton memoizes the calls which use $d$ in further derivations.
+When $a$ is deleted $d$ is stale, but as soon as $d$ is re-discovered, the any downstream computations memo-match and do not need to be re-computed.
+
+\subsection{Advantages}
+
+Datalog is implementation-agnostic with respect to the evaluation strategy for the engine itself. (Is this still true for IncA? I think so...)
+
+IncA likely outperforms a more generic strategy on edits that are friendly to it (monotonically increasing replacements).
 

### 2019-02-05 paper.tex
Added some citations to section 3

@@
 All of these are inadmissible for Flix or IncA.
+In the presence of non-monotone functions, IncA could not guarantee that monotonically increasing replacements would not require deletions.
+Therefore, to achieve from-scratch consistency, they would have to fallback to the naive incremental strategy.
 Adapton has no monotonicity restriction.\\
@@
 In contrast, Adapton memoizes the calls which use $d$ in further derivations.
-When $a$ is deleted $d$ is stale, but as soon as $d$ is re-discovered, the any downstream computations memo-match and do not need to be re-computed.
+When $a$ is deleted $d$ is stale, but as soon as $d$ is re-discovered, the any downstream computations memo-match and do not need to be re-computed.\\
+
+ There have been some attempts to alleviate the over-deletion problem (B/F alg \url{https://www.cs.ox.ac.uk/boris.motik/pubs/mnph15incremental-BF.pdf}) (hybrid alg \url{https://arxiv.org/pdf/1711.03987.pdf}).
+This appears to be an improvement to Delete/Rederive algorithms, but not in all cases, and Adapton should still give you this behavior effectively for free.\\
 

### 2019-02-19 paper.tex
Restructure w/explicit contributions + thread

@@
 
+We make the following contributions:
+
+\begin{enumerate}
+    \item We propose an efficient incremental naming strategy for the worklist algorithm over a CFG.
+    \item We provide an incremental abstract interpretation framework which can express any domain of interest. (Wording? Any domain + widening operator? Point is frees the monotonicity restriction)
+    \item We demonstrate that our framework is competitive, performance-wise, with 'Datalog-style' alternatives.
+\end{enumerate}
+
 \begin{figure}
@@
 
-Parity analysis (requires lattice)\\
-Interval analysis (requires widening)\\
-Octagon analysis (relational)
+Parity analysis (C1: demonstrates naming strategy)\\
+Interval analysis (C2: demonstrates extra expressivity)\\
+(C3 handled empirically in evaluation section)
 
-\section{Datalog Strategy - Limitations + Comparison}
+\section{Contribution: Naming Strategy for Worklist Algorithms}
 
-The Datalog-style approach to static analysis uses the following broad algorithm:
+We present our naming strategy for the generic worklist algorithm. You are convinced that this naming strategy gives some efficiency guarantee for an arbitrary analysis using this algorithm.\\
 
-\begin{enumerate}
-    \item Define a set of relations
-    \item Define a set of rules that refer to these relations
-    \item Transform the input program into a set of 'facts', which are judgements of membership in the relations
-    \item Apply the rules to generate new facts until a fixpoint is reached
-\end{enumerate}
+We contextualize our strategy via our simplest motivating example (parity).
 
-\subsection{Example}
+\section{Contribution: Expressive Abstract Interpretation Framework}
 
-As an example, consider the simple points-to analysis defined by the Flix authors (From Datalog to FLIX: A Declarative Language for Fixed Points on Lattices, section 2).
-The approach is applied as follows:
-
-\begin{enumerate}
-    \item Define a set of relations.\\
-    For this problem, there are 6 relations total - 4 "input relations", which are the base facts, and 2 "derived relations", which give the information we're interested in.
-They are as follows: New(v, h), Assign(v, v), Store(v, f, v), Load(v, v, f) are the inputs, and VarPointsTo(v, h) and HeapPointsTo(h, f, h) are the derived relations.
-Here the lowercase letters are types - v for variable, h for heap, f for field.
-    \newpage
-
-    \item Define a set of rules that refer to these relations.\\
-    There are four rules that define this analysis:
-    \begin{enumerate}
-        \item VarPointsTo(v1, h1) :- New(v1, h1)
-        \item VarPointsTo(v1, h2) :- Assign(v1, v2), VarPointsTo(v2, h2)
-        \item VarPointsTo(v1, h2) :- Load(v1, v2, f), VarPointsTo(v2, h1), HeapPointsTo(h1, f, h2)
-        \item HeapPointsTo(h1, f, h2) :- Store(v1, f, v2), VarPointsTo(v1, h1), VarPointsTo(v2, h2)
-    \end{enumerate}
-    For each of the rules, the fact on the left is added to the database if all of the facts on the right are in the database.\\
-
-    \item Transform the input program into a set of 'facts'.
-    \begin{lstlisting}[language=Java]
-    ClassA o1 = new ClassA(); // object A
-    ClassB o2 = new ClassB(); // object B
-    ClassB o3 = o2;
-    o2.f = o1;
-    Object r = o3.f;
-    \end{lstlisting}
-
-    This program is transformed into the following facts:\\
-    New("o1", "A")\\
-    New("o2", "B")\\
-    Assign("o3", "o2")\\
-    Store("o2", "f", "o1")\\
-    Load("r", "o3", "f")\\
-
-    The algorithm for this transformation is not given.
-For this simple example, it's fairly easy to see how each fact was derived from the program.
-Some algorithm for this transformation must be written for any analysis in this style.\\
-
-    \item Apply the rules to generate new facts until a fixpoint is reached.\\
-    This results in the following derived relations:\\
-    VarPointsTo("o1", "A") - rule 1\\
-    VarPointsTo("o2", "B") - rule 1\\
-    VarPointsTo("o3", "B") - rule 2\\
-    HeapPointsTo("B", "f", "A") - rule 4\\
-    VarPointsTo("r", "A") - rule 3\\
-
-    In the example, the final fact is what was desired.
-\end{enumerate}
-
-\subsection{Naive Incremental Computation in the Datalog World}
-
-In the Datalog context, an edit consists of the insertion or deletion of facts to the starting database.
-Wlog, we consider insertions and deletions of single facts and discuss naive incremental strategies for these edits.
-
-In the course of the Datalog engine's run, new facts are added to the database when a rule is triggered.
-It is easy to construct a simple 'provenance' of each fact in the database by tagging the fact with the facts that generated it.
-For instance, in the example from 3.1, the VarPointsTo("o3", "B") was generated via the facts Assign("o3", "o2") and VarPointsTo("o2, "B"), the latter of which was itself generated from New("o2", "B").
-Assuming that in the course of the algorithm these provenances are stored, consider now an edit which consists of the deletion of the fact New("o2", "B") from the initial database.
-Rather than re-computing from scratch, we can start from our finished database and do the following: delete the fact New("o2", "B"), delete all facts that transitively depend on New("o2", "B"), use the remaining facts as the initial database for a new run.
-Crucially, rather than re-computing all the facts from scratch, using this algorithm, any computed fact which does not depend on the deleted fact does not have to be re-computed.
-
-Now consider an edit which consists of the insertion of an arbitrary fact.
-This does not affect the facts in the database, which means we can do the following: simply use the finished database + the new fact as the initial database for a new run.
-But, what if the fact adds to a relation that is negated in one of our rules?
-Then the insertion of this fact is also a deletion of the corresponding negation, and the deletion should be processed first.
-(Datalog engines typically have restrictions on the use of negation in their rules - not discussed here in detail)
-
-Therefore, a naive incremental algorithm for an arbitrary set of deletions and insertions $\Delta\{d_k, i_k\}$ is as follows: we have a final database $D$.
-For each deletion $d_k$ , delete $d_k$ and all facts that transitively depend on it from $D$, leaving us with $D'$.
-Then add each inserted fact to get $D' + \{i_k\}$, and do a new run starting with this database.
-
-\subsection{IncA's Improved Approach}
-
-The above are the simplest two incremental strategies which many Datalog engines (even those that do not advertise themselves as incremental) implement.
-IncA, an explicitly incremental framework for Datalog (+ lattices), adds an additional optimization in the context of rules over lattices.
-Assuming all functions used are \textit{monotone}, a deletion of fact $d$ and insertion of fact $i$ where $d \sqsubseteq i$ constitutes a \textit{monotonically increasing replacement} (term introduced by IncA authors).
-If all functions used are monotone, then it is true that any fact derived from $d$ is also derivable from $i$ (proof left to IncA folks).
-Therefore, rather than deleting many facts derived from $d$ and subsequently re-deriving all of them, we don't have to perform any deletions.
-
-Therefore IncA's improved incremental algorithm is as follows.
-Given a final database $D$ and an arbitrary set of deletions and insertions $\Delta\{d_k, i_k\}$, first identify which $d_c$ have corresponding $i_c$ such that $d_c \sqsubseteq i_c$.
-Delete each of these from $\{d_k\}$ to get the strictly smaller set $\{d'_k\}$.
-Then run the above naive algorithm on $\Delta\{d'_k, i_k\}$
-
-There are some additional implementation details to IncA's algorithm that let it efficiently perform these deletions/re-derivations, but those do not change the improvement that's being made in terms of incremental computation, so we elide those details here.
-
-\subsection{Limitations}
-
-Once the Datalog strategy is chosen, there are several inherent limitations with respect to incremental computation.
-We briefly outline several of them here, and gesture toward how these are overcome using the Adapton strategy.\\
-
-Both Flix and IncA require that all functions over lattices are \textit{monotone}.
-This is a non-trivial restriction in the context of static analysis, as many standard analyses use non-monotonic widening functions to ensure termination.
-All of these are inadmissible for Flix or IncA.
-In the presence of non-monotone functions, IncA could not guarantee that monotonically increasing replacements would not require deletions.
-Therefore, to achieve from-scratch consistency, they would have to fallback to the naive incremental strategy.
-Adapton has no monotonicity restriction.\\
-
-There is an important sense in which the approach taken by IncA lacks 'history' in the sense of Adapton.
-We illustrate via an example.
-Consider an arbitrary lattice with elements $a, b, c, d$ such that the following holds: $a$ and $b$ are not ordered, $a \sqcup c = d$, $b \sqcup c = d$, $c \neq d$.
-Assuming that $a$ and $c$ are in the initial set of facts, an analysis run will add $d$ to the database.
-Now assume we make the following edit: delete $a$ and insert $b$.
-This is not a monotonically increasing replacement, so IncA will transitively delete all facts depending on $a$, including $d$ and any facts derived from $d$.
-Now after adding $b$, in the incremental run, the analysis will quickly re-discover $d$.
-However, since $d$ was already deleted from the database, all consequents of $d$ must be completely re-computed.
-Because IncA cannot 'look ahead' to see that $d$ will be re-discovered from an inserted fact, work that could have been saved is lost.
-This is a specific consequence of the general problem that deleting all facts deriving from a deleted fact is an over-approximation that is necessary for the Datalog approach to be sound.
-In contrast, Adapton memoizes the calls which use $d$ in further derivations.
-When $a$ is deleted $d$ is stale, but as soon as $d$ is re-discovered, the any downstream computations memo-match and do not need to be re-computed.\\
-
- There have been some attempts to alleviate the over-deletion problem (B/F alg \url{https://www.cs.ox.ac.uk/boris.motik/pubs/mnph15incremental-BF.pdf}) (hybrid alg \url{https://arxiv.org/pdf/1711.03987.pdf}).
-This appears to be an improvement to Delete/Rederive algorithms, but not in all cases, and Adapton should still give you this behavior effectively for free.\\
-
-\subsection{Advantages}
-
-Datalog is implementation-agnostic with respect to the evaluation strategy for the engine itself. (Is this still true for IncA? I think so...)
-
-IncA likely outperforms a more generic strategy on edits that are friendly to it (monotonically increasing replacements).
-
-\section{Contribution: Extending to Incremental Analysis Framework}
-
-We use the discussed data structures in a modular framework for implementing static analyses.
+We use the discussed algorithm in a modular framework for implementing abstract interpretation over arbitrary domains.
 Our framework requires the implementation of an abstract domain, including transfer function, to be given by the user.
@@
 Once we have defined the language to operate on - in our case, Jimple - and the transfer function for our domain of choice (see examples), we can incrementally execute the analysis.
-
-See figures 1/2 for an outline of the driving functions for the incremental analysis.
-By exploiting the structure of the typical 'worklist' algorithm for static analysis, we can achieve incremental performance gains.
-
-Our naming strategy works by reference to (insert defense of naming strategy - strictly location?) (reference to analysis trace visualization - given by Benno in table 3)
 
@@
 
-We evaluated our approach on a spread of analyses and target programs.
+Our evaluation seeks to answer several core research questions:
 
-Analyses:\\
-Interval, octagon.
+\begin{enumerate}
+    \item Does Patchwork achieve significant performance gains on incremental edits relative to a non-incremental baseline tool? (check overhead/speedup constants)
+    \item Is Patchwork competitive with alternative tools for incremental static analysis (IncA)? (direct comparison on parity analysis)
+    \item Is the added expressivity of Patchwork relevant to clients? (check false alarm rates on impoverished interval analysis v. regular)
+\end{enumerate}
 
-Programs:\\
-IncA benchmarks?
-
-Edit model:\\
-Adopt IncA approach?
-
-We compare to: IncA, off-the-shelf analysis?
-
-We show: same alarm rates as off-the-shelf analysis, similar incremental performance as IncA. QoIs: cumulative time, alarm rates, false alarm rates
-
-We need: off-the-shelf analysis to use, benchmarks to use, edit model to use
+We evaluated Patchwork on a series of benchmark programs (IncA's) using parity analysis, bad interval analysis, and normal interval analysis.
+We compare Patchwork's speed to Jandom on parity analysis (just that?).
+We compare Patchwork's speed to IncA on bad interval analysis.
+We compare the false alarm rate of the normal interval analysis to that of the bad interval analysis.
 
@@
 
-Interprocedural? Or doing that now?
+Interprocedural? Or doing that now?\\
+Improving naming strategy? Or specializing it to different domains?\\
+Incremental relation data structure?
 
@@
 \appendix
-\section{Appendix}
+\section{Appendix: Datalog Strategy - Limitations + Comparison}
 
-Text of appendix \ldots
+The Datalog-style approach to static analysis uses the following broad algorithm:
+
+\begin{enumerate}
+    \item Define a set of relations
+    \item Define a set of rules that refer to these relations
+    \item Transform the input program into a set of 'facts', which are judgements of membership in the relations
+    \item Apply the rules to generate new facts until a fixpoint is reached
+\end{enumerate}
+
+\subsection{Example}
+
+As an example, consider the simple points-to analysis defined by the Flix authors (From Datalog to FLIX: A Declarative Language for Fixed Points on Lattices, section 2).
+The approach is applied as follows:
+
+\begin{enumerate}
+    \item Define a set of relations.\\
+    For this problem, there are 6 relations total - 4 "input relations", which are the base facts, and 2 "derived relations", which give the information we're interested in.
+They are as follows: New(v, h), Assign(v, v), Store(v, f, v), Load(v, v, f) are the inputs, and VarPointsTo(v, h) and HeapPointsTo(h, f, h) are the derived relations.
+Here the lowercase letters are types - v for variable, h for heap, f for field.
+
+    \item Define a set of rules that refer to these relations.\\
+    There are four rules that define this analysis:
+    \begin{enumerate}
+        \item VarPointsTo(v1, h1) :- New(v1, h1)
+        \item VarPointsTo(v1, h2) :- Assign(v1, v2), VarPointsTo(v2, h2)
+        \item VarPointsTo(v1, h2) :- Load(v1, v2, f), VarPointsTo(v2, h1), HeapPointsTo(h1, f, h2)
+        \item HeapPointsTo(h1, f, h2) :- Store(v1, f, v2), VarPointsTo(v1, h1), VarPointsTo(v2, h2)
+    \end{enumerate}
+    For each of the rules, the fact on the left is added to the database if all of the facts on the right are in the database.\\
+
+    \item Transform the input program into a set of 'facts'.
+    \begin{lstlisting}[language=Java]
+    ClassA o1 = new ClassA(); // object A
+    ClassB o2 = new ClassB(); // object B
+    ClassB o3 = o2;
+    o2.f = o1;
+    Object r = o3.f;
+    \end{lstlisting}
+
+    This program is transformed into the following facts:\\
+    New("o1", "A")\\
+    New("o2", "B")\\
+    Assign("o3", "o2")\\
+    Store("o2", "f", "o1")\\
+    Load("r", "o3", "f")\\
+
+    The algorithm for this transformation is not given.
+For this simple example, it's fairly easy to see how each fact was derived from the program.
+Some algorithm for this transformation must be written for any analysis in this style.\\
+
+    \item Apply the rules to generate new facts until a fixpoint is reached.\\
+    This results in the following derived relations:\\
+    VarPointsTo("o1", "A") - rule 1\\
+    VarPointsTo("o2", "B") - rule 1\\
+    VarPointsTo("o3", "B") - rule 2\\
+    HeapPointsTo("B", "f", "A") - rule 4\\
+    VarPointsTo("r", "A") - rule 3\\
+
+    In the example, the final fact is what was desired.
+\end{enumerate}
+
+\subsection{Naive Incremental Computation in the Datalog World}
+
+In the Datalog context, an edit consists of the insertion or deletion of facts to the starting database.
+Wlog, we consider insertions and deletions of single facts and discuss naive incremental strategies for these edits.
+
+In the course of the Datalog engine's run, new facts are added to the database when a rule is triggered.
+It is easy to construct a simple 'provenance' of each fact in the database by tagging the fact with the facts that generated it.
+For instance, in the example from 3.1, the VarPointsTo("o3", "B") was generated via the facts Assign("o3", "o2") and VarPointsTo("o2, "B"), the latter of which was itself generated from New("o2", "B").
+Assuming that in the course of the algorithm these provenances are stored, consider now an edit which consists of the deletion of the fact New("o2", "B") from the initial database.
+Rather than re-computing from scratch, we can start from our finished database and do the following: delete the fact New("o2", "B"), delete all facts that transitively depend on New("o2", "B"), use the remaining facts as the initial database for a new run.
+Crucially, rather than re-computing all the facts from scratch, using this algorithm, any computed fact which does not depend on the deleted fact does not have to be re-computed.
+
+Now consider an edit which consists of the insertion of an arbitrary fact.
+This does not affect the facts in the database, which means we can do the following: simply use the finished database + the new fact as the initial database for a new run.
+But, what if the fact adds to a relation that is negated in one of our rules?
+Then the insertion of this fact is also a deletion of the corresponding negation, and the deletion should be processed first.
+(Datalog engines typically have restrictions on the use of negation in their rules - not discussed here in detail)
+
+Therefore, a naive incremental algorithm for an arbitrary set of deletions and insertions $\Delta\{d_k, i_k\}$ is as follows: we have a final database $D$.
+For each deletion $d_k$ , delete $d_k$ and all facts that transitively depend on it from $D$, leaving us with $D'$.
+Then add each inserted fact to get $D' + \{i_k\}$, and do a new run starting with this database.
+
+\subsection{IncA's Improved Approach}
+
+The above are the simplest two incremental strategies which many Datalog engines (even those that do not advertise themselves as incremental) implement.
+IncA, an explicitly incremental framework for Datalog (+ lattices), adds an additional optimization in the context of rules over lattices.
+Assuming all functions used are \textit{monotone}, a deletion of fact $d$ and insertion of fact $i$ where $d \sqsubseteq i$ constitutes a \textit{monotonically increasing replacement} (term introduced by IncA authors).
+If all functions used are monotone, then it is true that any fact derived from $d$ is also derivable from $i$ (proof left to IncA folks).
+Therefore, rather than deleting many facts derived from $d$ and subsequently re-deriving all of them, we don't have to perform any deletions.
+
+Therefore IncA's improved incremental algorithm is as follows.
+Given a final database $D$ and an arbitrary set of deletions and insertions $\Delta\{d_k, i_k\}$, first identify which $d_c$ have corresponding $i_c$ such that $d_c \sqsubseteq i_c$.
+Delete each of these from $\{d_k\}$ to get the strictly smaller set $\{d'_k\}$.
+Then run the above naive algorithm on $\Delta\{d'_k, i_k\}$
+
+There are some additional implementation details to IncA's algorithm that let it efficiently perform these deletions/re-derivations, but those do not change the improvement that's being made in terms of incremental computation, so we elide those details here.
+
+\subsection{Limitations}
+
+Once the Datalog strategy is chosen, there are several inherent limitations with respect to incremental computation.
+We briefly outline several of them here, and gesture toward how these are overcome using the Adapton strategy.\\
+
+Choosing the Datalog strategy does not allow the user to directly reference the program of interest in their analysis.
+Instead, they must 'compile' the program into a new set of facts/relations relevant to their analysis.
+Often, as in IncA's interval analysis, a considerable amount of scaffolding relations are necessary to retain awareness of the program's control flow, and the user must correctly connect the program's semantics to the relations and types written for Datalog, which requires consideration of two distinct domains.
+In contrast, the Adapton approach allows us to directly write abstract transfer functions over the concrete semantics of our language of choice, removing a layer of abstraction.\\
+
+Both Flix and IncA require that all functions over lattices are \textit{monotone}.
+This is a non-trivial restriction in the context of static analysis, as many standard analyses use non-monotonic widening functions to ensure termination.
+All of these are inadmissible for Flix or IncA.
+In the presence of non-monotone functions, IncA could not guarantee that monotonically increasing replacements would not require deletions.
+Therefore, to achieve from-scratch consistency, they would have to fallback to the naive incremental strategy.
+Adapton has no monotonicity restriction.\\
+
+There is an important sense in which the approach taken by IncA lacks 'history' in the sense of Adapton.
+We illustrate via an example.
+Consider an arbitrary lattice with elements $a, b, c, d$ such that the following holds: $a$ and $b$ are not ordered, $a \sqcup c = d$, $b \sqcup c = d$, $c \neq d$.
+Assuming that $a$ and $c$ are in the initial set of facts, an analysis run will add $d$ to the database.
+Now assume we make the following edit: delete $a$ and insert $b$.
+This is not a monotonically increasing replacement, so IncA will transitively delete all facts depending on $a$, including $d$ and any facts derived from $d$.
+Now after adding $b$, in the incremental run, the analysis will quickly re-discover $d$.
+However, since $d$ was already deleted from the database, all consequents of $d$ must be completely re-computed.
+Because IncA cannot 'look ahead' to see that $d$ will be re-discovered from an inserted fact, work that could have been saved is lost.
+This is a specific consequence of the general problem that deleting all facts deriving from a deleted fact is an over-approximation that is necessary for the Datalog approach to be sound.
+In contrast, Adapton memoizes the calls which use $d$ in further derivations.
+When $a$ is deleted $d$ is stale, but as soon as $d$ is re-discovered, the any downstream computations memo-match and do not need to be re-computed.\\
+
+ There have been some attempts to alleviate the over-deletion problem (B/F alg \url{https://www.cs.ox.ac.uk/boris.motik/pubs/mnph15incremental-BF.pdf}) (hybrid alg \url{https://arxiv.org/pdf/1711.03987.pdf}).
+This appears to be an improvement to Delete/Rederive algorithms, but not in all cases, and Adapton should still give you this behavior effectively for free.\\
+
+\subsection{Advantages}
+
+Datalog is implementation-agnostic with respect to the evaluation strategy for the engine itself. (Is this still true for IncA? I think so...)
+
+IncA likely outperforms a more generic strategy on edits that are friendly to it (monotonically increasing replacements).
 

### 2019-02-25 paper.tex
Flesh out contributions, add REVISER section

@@
 \begin{enumerate}
-    \item We propose an efficient incremental naming strategy for the worklist algorithm over a CFG.
-    \item We provide an incremental abstract interpretation framework which can express any domain of interest. (Wording? Any domain + widening operator? Point is frees the monotonicity restriction)
-    \item We demonstrate that our framework is competitive, performance-wise, with 'Datalog-style' alternatives.
+    \item We propose an efficient incremental naming strategy for the worklist algorithm over a CFG.\\
+    Using Adapton requires a strategy for uniquely naming subcomputations inside an algorithm.
+The goal is to ensure that after a minor edit, identical subcomputations have the same name across runs, so that Adapton can re-use work from a prior run.
+We present a naming strategy which names the transfer function at each CFG node by the program location and current abstract state.
+We demonstrate that this achieves significant re-use across runs in the presence of typical small edits to a program.
+    \item We provide an incremental abstract interpretation framework which can express any domain of interest.\\
+    Patchwork allows clients to use our naming strategy for arbitrary abstract domains.
+The user must supply a language specification and an implementation of the desired transfer function.
+With this supplied, Patchwork runs the analysis and can execute incremental runs.
+We demonstrate the use of Patchwork on several numerical domains - parity and interval - but non-numerical domains can be used as well.
+There are no restrictions on the domains or transfer functions supplied by the user, an improvement over alternatives which require all operators to be monotonic.
+    \item We demonstrate that our framework is competitive, performance-wise, with 'Datalog-style' alternatives.\\
 \end{enumerate}
@@
     \item Does Patchwork achieve significant performance gains on incremental edits relative to a non-incremental baseline tool? (check overhead/speedup constants)
+    \item Is the added expressivity of Patchwork relevant to clients? (check false alarm rates on impoverished interval analysis v. regular)
     \item Is Patchwork competitive with alternative tools for incremental static analysis (IncA)? (direct comparison on parity analysis)
-    \item Is the added expressivity of Patchwork relevant to clients? (check false alarm rates on impoverished interval analysis v. regular)
 \end{enumerate}
@@
 For instance, IncA's framework could not express (our worked example) - they would instead have to use (different widening operator), which would significantly degrade the strength of the analysis (for reasons).
+
+Reviser: Reviser is similar to Patchwork in that it tracks the graph structure of static analyses to achieve incremental performance games.
+Reviser uses the IDFS framework, which builds an exploded CFG tracking both control flow and abstract state dependencies in the program.
+Like Adapton, Reviser identifies portions of the computation changed by an edit and only 'cleans' the graph edges reachable from the edited regions.
+However, Reviser is tightly coupled with the IDFS framework, and cannot express more general graph-based computations.
+Moreover, the performance gains from Reviser are entirely from re-computing over a subset of the original graph - the creators indicate that their clean/repropogate algorithm is \textit{slower} than an initial run - as opposed to Adapton's ability to cache and re-use memoized data.
 
@@
 
+\newpage
 %% Appendix
